---
title: "Effects of Class Size on Math Performance in Tennessee"
output:
  html_document:
    df_print: paged
  pdf_document:
    df_print: paged
    fig_caption: yes
    number_sections: yes
---

<style type="text/css">

body{ /* Normal  */
      font-size: 18px;
  }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE)
```

```{r, include=FALSE}
library("AER")
library("foreign")
library("sqldf")
library("plotly")
library("kableExtra")
library("MASS")
library("ggplot2")
```


```{r, include=FALSE}
setwd("C:\\Users\\rosen_000\\Dropbox\\Daivs\\2-Winter Q 2020\\STA 207 Research Methods II\\Project-2")
star = read.spss("STAR_Students.sav",to.data.frame = TRUE)
```


```{r, include = FALSE}
data1 = names(star) %in% c("g1tchid","g1classtype","g1schid","g1surban", "g1classsize","g1tmathss", "g1tgen")
data1 = star[,data1]
data1 = data1[complete.cases(data1),]
```





```{r, include = FALSE}
#Group by teachers as unit
t_data1 = sqldf("SELECT DISTINCT(g1tchid) as g1_teachID,
                            g1schid as g1_schoolID,
                            g1classtype as g1_classtype,
                            g1surban as g1_surburban,
                            COUNT(g1tchid) as g1_classsize,
                            AVG(g1tmathss) as g1_avgmath,
                            MIN(g1tmathss) as g1_minmath,
                            MAX(g1tmathss) as g1_maxmath,
                            MEDIAN(g1tmathss) as g1_medmath
                    FROM data1
                    GROUP BY g1tchid")


t_data1$g1_teachID = as.factor(t_data1$g1_teachID)
t_data1$g1_schoolID = as.factor(t_data1$g1_schoolID)
t_data1$g1_classtype= as.factor(t_data1$g1_classtype)
t_data1$g1_surburban = as.factor(t_data1$g1_surburban)



#write.csv(t_data1, "Teach_as_Unit.csv")
```

```{r, include = FALSE}
#Boxplot of average class math scores broken by class type
box1 = plot_ly(t_data1, y = ~g1_avgmath, color = ~g1_classtype, type = "box") %>%
  layout(title = "Average Classroom Score",
         yaxis = list(title = "Classroom Average Sclaed Math Score"),
         xaxis = list(title = "Class Type"))

#Boxplot of min class math scores broken by class type
box2 = plot_ly(t_data1, y = ~g1_minmath, color = ~g1_classtype, type = "box") %>%
  layout(title = "Minimum Classroom Score",
         yaxis = list(title = "Classroom Minimum Sclaed Math Score"),
         xaxis = list(title = "Class Type")) 

#Boxplot of min class math scores broken by class type
box3 = plot_ly(t_data1, y = ~g1_maxmath, color = ~g1_classtype, type = "box") %>%
  layout(title = "Classroom Maximum Score",
         yaxis = list(title = "Sclaed Math Score"),
         xaxis = list(title = "Class Type"))

#Boxplot of med class math scores broken by class type
box4 = plot_ly(t_data1, y = ~g1_medmath, color = ~g1_classtype, type = "box") %>%
  layout(title = "Classroom Median Score",
         yaxis = list(title = "Sclaed Math Score"),
         xaxis = list(title = "Class Type"))

#Boxplot of med class math scores broken by schoolID
box5 = plot_ly(t_data1, y = ~g1_medmath, color = ~g1_schoolID, type = "box") %>%
  layout(title = "Classroom Median Score",
         yaxis = list(title = "Sclaed Math Score"),
         xaxis = list(title = "SchoolID"))
```


```{r, include = FALSE}
#Histograms for Average Scores
a = sqldf("SELECT g1_medmath FROM t_data1
          WHERE g1_classtype == 'SMALL CLASS'")

b = sqldf("SELECT g1_medmath FROM t_data1
          WHERE g1_classtype  == 'REGULAR + AIDE CLASS'")

c = sqldf("SELECT g1_medmath FROM t_data1
          WHERE g1_classtype == 'REGULAR CLASS'")

hist_avgsmall <- plot_ly(alpha = 0.2) %>%
  add_histogram(x = ~a$g1_medmath, marker = list(color = "#C6CFE5")) %>%
  layout(barmode = "overlay",
         title = "Class Median Scaled Math Score",
         xaxis = list(title = "SMALL CLASSES"))

hist_avgreg <- plot_ly(alpha = 0.2) %>%
  add_histogram(x = ~b$g1_medmath, marker = list(color = "#FDC6B0")) %>%
  layout(barmode = "overlay",
         title = "Class Median Scaled Math Score",
         xaxis = list(title = "REGULAR CLASSES"))

hist_avgregaide <- plot_ly(alpha = 0.2) %>%
  add_histogram(x = ~c$g1_medmath, marker = list(color = "#B2E0D2")) %>%
  layout(barmode = "overlay",
         title = "Class Median Scaled Math Score",
         xaxis = list(title = "REGULAR + AIDE CLASSES"))

```


```{r, include = FALSE}
Lv_test_med1 = leveneTest(g1_avgmath~g1_classtype, data = t_data1) #Equal Variacnes
kable(Lv_test_med1, caption = "Levene's Test for Equal Variance: Classroom Median Score") %>% kable_styling(bootstrap_options = "striped", full_width = F)


Lv_test_med2 = leveneTest(g1_avgmath~g1_schoolID, data = t_data1) #Equal Variacnes
kable(Lv_test_med2, caption = "Levene's Test for Equal Variance: SchoolID") %>% kable_styling(bootstrap_options = "striped", full_width = F)
```



***

### Team ID: Team 6

#### NAME: Connor Rosenberg
#### NAME: Rongkui Han
#### NAME: Yuqing Yang
#### NAME: Nassim Ali-Chaouche

***

## 1.0 Introduction


### 1.1 Background

The Student/Teacher Achievement Ratio (STAR) was a four-year longitudinal class-size study funded by the Tennessee General Assembly and conducted by the State Department of Education. Over 7,000 students from kindergarten to 3rd grade in 79 schools were randomly assigned into one of three interventions: small class (13 to 17 students per teacher), regular class (22 to 25 students per teacher), and regular-with-aide class (22 to 25 students with a full-time teacher's aide). The interventions were initiated as the students entered school in kindergarten and continued through third grade.     

Besides following standard procedures that ensured confidentiality and ethics in human subjects' research, Project STAR also highlights the important features of:    

1. *Each school included in the study had to have a large enough student body to form as least one of each of the three class types.*  

2. *Students and teachers were randomly assigned to their class type.*    

Project STAR is an example of **stratified randomized design**, where experimental units are grouped together according to certain pre-treatment characteristics into strata. Within each stratum, a completely randomized experiment is conducted. In the case that there exist population structures that associate or covary with the experimental outcome, stratified randomized experiments generally are more informative than completely randomized experiments (Suresh K., 2011). The goal of a stratified study is usually not to identify treatment effects within a single stratum, but rather treatment effects across all strata.

Because it is reasonable to expect systemic differences in educational outcome across schools, due to reasons such as demographics, school could be a confounding variable and influence the ooutcome of the Project STAR research. Therefore, each school should be viewed as a stratum in analyzing Project STAR data.   

To expand on previous findings, teachers, rather than individual student, will be used as experimental units. The adjustment of experimental unit will facilitate the making of a causal statement as to the effect of class size on educational outcome, because the teacher experimental units more convincingly satisfy the stable unit treatment value assumption (SUTVA) and independence assumptions necessary for causal inferences. The detail of these assumtions will be dissected in the discussion section. 

### 1.2 Questions of Interest

Our questions of interest are as follows:     

1. Is there a significant difference in a first-grade teacher's teaching performance in math across the three different class sizes?    
2. Are teacher's performances relatively stable between different schools? That is, does the school itself affect class average math scores?    
3. Does our ANOVA model fit well with the data? In other words, are the analysis of variance assumptions satisfied?    
4. Can we draw causal conclusion that class sizes affect the class average math scores of first-grade teachers?       

## 2.0 Analysis Plan

### 2.1 Population and study design

Project STAR is an example of stratified randomized design, where experimental units are grouped together according to certain pre-treatment characteristics. Within each stratum, a randomized experiment is conducted. In this study, each school can be viewed as a stratum and each of the three different types of class sizes are the experiment. A two-way ANOVA test is fitting to answer our questions of interest under this study design. One factor in the ANOVA model will be class size treatment, whose main effect on teacher performance is of primary interest to us. The other factor will be school ID, in order to control for and observe the stratum effect. 

To expand on our previous findings, we will set teachers as the experimental unit, rather than the individual student. We will use the *median* scaled 1st grade math score of all students under each teacher for our analysis. The median score of a class more faithfully reflects the class' performance under each teacher. In additon, the median is usually a more robust summary statistic than the mean, because it is less affected by outliers. The adjustment of experimental unit will enable us to make a causal statement as to the effect of class size on educational outcome.

A two-way ANOVA test is fitting for answering our questions of interest under the stratified randomized design. One factor in the ANOVA model will be class size, whose main effect is of primary interest in this study. The other factor will be school ID, in order to control for and observe the stratum effect. 

In oreder to treat teachers as the experimental units, we will use the *median* scaled 1st grade math score of all students under each teacher for our analysis. The median score of a class truthfully reflects the class' performance. In additon, the median is usually a more robust summary statistic than the mean, because it is less affected by outliers. 

### 2.2 Statistical Analysis

#### 2.2.1 Descriptive Analysis

We will check 
- Completedness and balance of each stratum; 
- Variation of response variable, scaled 1st grade math score, across class types; 
- Variation of response variable across strata;


Figure .. completedness of strata

```{r}
t_data1_tally = t_data1 %>%
  group_by(g1_schoolID, g1_classtype) %>%
  tally()
ggplot(t_data1_tally, aes(x = g1_schoolID, y = n, fill = g1_classtype)) +
  geom_col()
```

#### 2.2.2 Main Analysis

For our main analysis, we will construct the following factor effects model for the classroom median math score.

$$y_{ijk} = μ_{..} + τ_i + β_j + ε_{ ijk}$$
for $i$ $\epsilon$ $[1,2,3,4]$ , $j$ $\epsilon$ $[1,2,...,76]$ , and $k$ $\epsilon$ $[1,2,...,338]$

Our model is constrained such that $\sum_{i=1}^{4} τ_i = 0$ and $\sum_{i=1}^{76} β_j = 0$. These two constraints imply the following impact on our parameter interpretations. 

- $μ_{..}$ represents the overall classroom median score across all treatment levels.   
- $τ_i$ represents the effect of each class size on the overall median math score.   
- $β_j$ represents the effect of each school on the overall median math score.


Our model is appropriate to answer the questions of interest because it captures the effect of each treatment on the class’ median performance while controlling for other external factors by blocking by school id.

Figure #, which shows the boxplots of the classroom median scores for each school, highlights the variability in teacher performance across each distinct school. This variability is likely due to the similar demographic features within schools, but various demographic features between them. For example, schools located in areas of high affluence may achieve better classroom performance since more students have access to academic support in addition to greater parent oversight. Similarly, schools who pull their students from less affluent areas may see worse classroom performance due to student food insecurity, lack of academic support, and other social deficiencies.

```{r, echo=FALSE}
box5
```

Because of this high variability in median classroom performance between schools, blocking by school ID in our model helps to extract the true effect of the class size treatment on the classroom median math performance. 


From the model equation above, you can see that we are not considering the interaction effect between class type and school id.  We made this decision for two reasons. First, the inclusion of an interaction term would increase the number of parameters which need estimation from 79 to a number which exceeds our sample size. This would impede us from fitting a proper model and is our first rationale for omitting this term in our analysis. Our second reason is that of relevance. The purpose of our analysis is to examine the main effects of class size on the median scaled math score achieved by a first-grade teacher’s class. We want out conclusions to be applicable to the most general a population possible. If we include an interaction term, our model now describes the effect of class sizes within each individual school & teacher combination. For the combination of these two arguments, we will omit the interaction parameter between class size and school id from our model. 
 

### 2.3 Sensitivity Analysis

> Talk about how we will conduct the sensitivity analysis

## 3.0 Results

### 3.1 Descriptive Analysis

### 3.2 Main Analysis


```{r, echo=FALSE}
med_anova = aov(g1_medmath ~ g1_classtype + g1_schoolID, 
                  data = t_data1)

med_anova1 = anova(lm(g1_medmath ~ g1_classtype + g1_schoolID,
                          data = t_data1))


med_anova1 = round(med_anova1, 2)
med_anova1$`Pr(>F)` = c("<.001", "<.001", "NA")


kable(med_anova1, caption  = "Model ANOVA Table") %>% kable_styling(bootstrap_options = "striped", full_width = F)
```


> Task 6: Test whether there is a difference in math scaled score in 1st grade across teachers in different class types. Justify your choice of test.

### 3.3 Sensitivity analysis

#### 3.3.1 Normality
Figure 

From the Q-Q plot of the residuals (Figure ..) , we can observe that most of that data points lie on a straight line, which is close to what we expect to see from a normal distribution. There are some points at the right tail which have a higher probability mass than expected; however, these points are only a few compared to the total amount of data. Thus, the normality assumption is largely satisfied from the Q-Q plot. 

```{r, echo = FALSE, fig.align="center"}
plot(med_anova, which = c(2))
hist(reg1$residuals, main = "Histogram of Residuals", xlab = "Residual Values")

```



To visualize the distribution of the residuals, a histogram will be used:

From the histogram of the residuals, we can observe that the distribution of the residuals looks symmetric about the mean and bell-shaped, similar to what is seen from a normal distribution. 

To further test for normality of the errors, a Shapiro-Wilk test will be used on the distribution of the residuals. A Shapiro-Wilk test is used to test whether a distribution of data follows a normal distribution.

The hypotheses are as follows:
$H_o$: The residuals are normally distributed.
$H_a$: The residuals are not normally distributed.

```{r, echo = FALSE}

W.statistic = round(as.numeric(names(table(shapiro.test(reg1$residuals)$statistic))), 2)
P.value = round(shapiro.test(reg1$residuals)$p.value, 2)

table = cbind(W.statistic, P.value)
colnames(table) = c('W Statistic', 'P-Value')
kable(table, caption = "Shapiro-Wilk Normality Test") %>% kable_styling(bootstrap_options = c("striped", "bordered"), full_width = F)


```


The p-value of 0.13 is greater than the significance level of 0.05, and thus we fail to reject the null hypothesis. Thus, there is no evidence that the distribution of the residuals does not follow a normal distribution. 




#### 3.3.2 Equal Variances:

Residual vs Fitted Plot:

```{r, echo = FALSE, fig.align="center"}
plot(med_anova, which = c(1))
```


The residuals are evenly spread out along the x-axis, and thus it seems that the equal variance assumption is satisfied.

A Levene's test will be used to further test the equal variance assumption for both variables. 

The hypotheses are as follows:
$H_o$: The variances are equal across the various groups.
$H_a$: Not all variances are equal across groups.

```{r, echo = FALSE}
a = leveneTest(g1_medmath~g1_classtype, data = t_data1)$Df[1]
b = leveneTest(g1_medmath~g1_classtype, data = t_data1)$Df[2]
c = paste(a, b, sep = ", ")
d = round(leveneTest(g1_medmath~g1_classtype, data = t_data1)$F[1], 2)
e = round(leveneTest(g1_medmath~g1_classtype, data = t_data1)$P[1], 2)

table1 = cbind(c, d, e)
colnames(table1) = c('df', 'F Value', 'Pr(>F)')
kable(table1, caption = "Levene Test for Class Type Variable") %>% kable_styling(bootstrap_options = c("striped", "bordered"), full_width = F)

```


```{r, echo = FALSE}
a = leveneTest(g1_medmath~g1_schoolID, data = t_data1)$Df[1]
b = leveneTest(g1_medmath~g1_schoolID, data = t_data1)$Df[2]
c = paste(a, b, sep = ", ")
d = round(leveneTest(g1_medmath~g1_schoolID, data = t_data1)$F[1], 2)
e = round(leveneTest(g1_medmath~g1_schoolID, data = t_data1)$P[1], 2)

table1 = cbind(c, d, e)
colnames(table1) = c('df', 'F Value', 'Pr(>F)')
kable(table1, caption = "Levene Test for School ID Variable") %>% kable_styling(bootstrap_options = c("striped", "bordered"), full_width = F)
```


In the Levene's test for both variables, both p-values are greater than the significance level of 0.05. Thus for each variable, there is no evidence that not all of the variances are equal across treatment groups. Thus, the equal variance assumption is satisfied.  






#### 3.3.3 Independence:

This experiment is randomized in two ways and represents the best controlled environment to achieve independence of the results. First of all, teachers are randomly assigned to each class type: small, regular, and regular with aide. Second, every student is randomly assigned to each teacher. There is a possibility of the independence assumption not holding completely from the time of randomization to the time that the test scores were recorded. For instance, teachers may share materials, or parents may seek out tutoring for their child. However, given the randomized design of the experiment and the relatively large sample size, we may assume that the results are essentially independent. 

#### 3.3.4 Outliers:

An outlier is defined as an observation whose studentized residual has a magnitude greater than three.

```{r, echo = FALSE}

a = length(which(studres(reg1)>3 | studres(reg1)< -3))
b = names(which(studres(reg1)>3 | studres(reg1)< -3))

table2 = cbind(a, b)
colnames(table2) = c('Number of Outliers', 'Observation Numbers')
kable(table2, caption = "Outliers") %>% kable_styling(bootstrap_options = c("striped", "bordered"), full_width = F)

```

There is only one outlier in the data, which is observation number 252. Thus, outliers do not pose a problem in our analysis.



## 4.0 Discussion

In this report, we presented our usage of 2-way ANOVA to analyze the effect of class size on first-grade teachers' teaching performance in math in a stratified randomized experiment, using each school as a stratum. We explored the effect of including schoo-by-class size interactions in our model, and concluded that interactions between the two factors did not contribute significantly to the variance partitioning of the data. Model diagnostics suggested that the dataset satisfied assumptions for ANOVA. Results derived from the fitting of the model suggested significant difference in a first-grade teacher's median math scores across different class sizes. Pairwise comparisons suggested that **small classes and regular classes with aides both outperformed regular classes without aides**. The model also revealed significant performance differences across schools, with the largest pairwaise difference being **XXX** in class median 1st grade scale math score.      

This analysis enables us to make causal statements regarding the effect of class size on teacher's performance in math education. This is made possible by using teachers as experimental units, thus satisfying the SUTVA and independence assumptions necessary for causal inferences:    

**SUTVA**:
Definition: *The potential outcomes for any unit do not vary with the treatments assigned to other units, and, for each unit, there are no different forms or versions of each treatment level, which lead to different potential outcomes.*     
The experimental unit used in the analysis first satisfies the no-interference component of SUTVA – the assumption that the treatment
applied to one unit does not affect the outcome for other units. On the basis of prior knowledge of school systems, it is realistic to assume that one teacher being assigned to a specific class size does not affect the teaching outcoome of another teacher. The second component of SUTVA requires that individuals receiving a specific treatment cannot receive different forms of that treatment. In our case, due to the strict randomization implemented in the experiment, the class taught by one teacher is by nature homogenous with a class taught by another.       

**Independence Assumption**:    
Definition: *the assignment of treatment is independent of potential outcomes of experimental units.*   
This assumption is met in the experiment by using double randomization: One random assignment is that of teachers to classes. The second randomization is of students to classes/teachers. The design ensures that high/low performance teacher or students were not systematically enriched in any class-size treatments. In light of this, systematic effects can be interpreted as the effects of class size. 

Therefore, our analysis concludes that smaller class size has a positive average causal effect on a teacher's teaching outcome in math. 
This is different from the conclusion of Project I. SUTVA was not plausible when using individual students as experimental units. Interactions between students likely resulted in altered potential outcome of one student due to the treatment assigned to another, thus violating SUTVA. In that case, rejections of the null hypothesis would not necessarily be convincing evidence of effects of class size; it may simply indicate the presence of peer effects. In contrast, using teachers as experimental units does not rely on no-interference assumptions among students. This makes the results reported here credible evidence of causal class-size effects.

## 5.0 Reference
Suresh K. (2011). An overview of randomization techniques: An unbiased assessment of outcome in clinical research. Journal of human reproductive sciences, 4(1), 8–11. doi:10.4103/0974-1208.82352    



***
***
