---
title: "Effects of Class Size on Math Performance in Tennessee"
output:
  pdf_document:
    df_print: paged
    fig_caption: yes
    number_sections: yes
---

<style type="text/css">

body{ /* Normal  */
      font-size: 18px;
  }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE)
```

```{r}
library("AER")
library("foreign")
library("MASS")
library("kableExtra")

```


```{r}
setwd("~/Desktop/STA 207/Project 2/Project 2 Github/Project-2")
star = read.spss("STAR_Students.sav",to.data.frame = TRUE)







```

```{r}
data = names(star) %in% c("g1tchid","g1classtype","g1schid","g1surban", "g1classsize", "g1mathss", "g1tgen","g1trace","g1thighdegree","g1tcareer","g1tyears","g1motivraw","g1selfconraw", "g2tchid","g2classtype","g2schid","g2surban", "g2classsize", "g2absent","g2mathss", "g2tgen","g2trace","g2thighdegree","g2tcareer","g2tyears","g2motivraw","g2selfconraw")
data = star[,data]
data = data[complete.cases(data),]
```

```{r}
data1 = names(star) %in% c("g1tchid","g1classtype","g1schid","g1surban", "g1classsize","g1mathss", "g1tgen","g1trace","g1thighdegree","g1tcareer","g1tyears","g1motivraw","g1selfconraw")
data1 = star[,data1]
data1 = data1[complete.cases(data1),]
```

```{r}
data2 = names(star) %in% c("g2tchid","g2classtype","g2schid","g2surban", "g2classsize", "g2absent","g2mathss", "g2tgen","g2trace","g2thighdegree","g2tcareer","g2tyears","g2motivraw","g2selfconraw")
data2 = star[,data2]
data2 = data2[complete.cases(data2),]
```




```{r, include=FALSE}
#Write Report
```

***

Team ID:

Name:
Name:
Name:
Name:



# Introduction


## Background

The Student/Teacher Achievement Ratio (STAR) was a four-year longitudinal class-size study funded by the Tennessee General Assembly and conducted by the State Department of Education. Over 7,000 students from kindergarten to 3rd grade in 79 schools were randomly assigned into one of three interventions: small class (13 to 17 students per teacher), regular class (22 to 25 students per teacher), and regular-with-aide class (22 to 25 students with a full-time teacher's aide). Classroom teachers were also randomly assigned to the classes they would teach. The interventions were initiated as the students entered school in kindergarten and continued through third grade.    

The STAR experiment was designed by a group of researchers including Helen Pate-Bain, the driving force behind Project STAR, other academics, and members of the Tennessee Department of Education. Some of its key features are: 

1. *All Tennessee schools with K-3 classes were invited to participate.* Giving every school a chance to join the study helped ensure a diverse sample as well as rule out the possibility that class-size effects could be attributed to selection bias.   

2. *Each school included in the study had to have a large enough student body to form as least one of each of the three class types.* The within-school design provided built-in control for differences among schools in terms of resources, leadership, and facilities.   

3. *Schools from inner-city, urban, suburban and rural locations were included in the experiment.* This feature guaranteed that samples would include children from various ethnic backgrounds and income levels.   

4. *Students and teachers were randomly assigned to their class type.*    

5. *Investigators followed the standard procedures for confidentiality in human subjects' research.*   

6. *No children were to receive fewer services than normal because of the experiment.*    

7. *Student achievement was to be tracked by standarized tests, which were carefully monitored.*   


## Statistical questions of interest


# Analysis Plan


## Population and study design


## Statistical Analysis

### Descriptive Analysis

### Main Analysis

### Sensitivity Analysis

# Results

## Descriptive Analysis


## Inferential Analysis

## Sensitivity analysis


# Discussion



# Model Diagnostics


```{r, include = FALSE}
Teach_as_Unit <- read.csv("~/Desktop/STA 207/Project 2/Project 2 Github/Project-2/Teach_as_Unit.csv")
t_data1 = Teach_as_Unit
```


```{r, include = FALSE}

t_data1$g1_schoolID = as.factor(t_data1$g1_schoolID)

reg1 = lm(g1_medmath ~ g1_classtype + g1_schoolID, 
                  data = t_data1)

med_anova = aov(g1_medmath ~ as.factor(g1_classtype) + as.factor(g1_schoolID), 
                  data = t_data1)

```




# Normality:

```{r, echo = FALSE, fig.align="center"}
plot(med_anova, which = c(2))
```
Figure 

From the Q-Q plot of the residuals (Figure ..) , we can observe that most of that data points lie on a straight line, which is close to what we expect to see from a normal distribution. There are some points at the right tail which have a higher probability mass than expected; however, these points are only a few compared to the total amount of data. Thus, the normality assumption is largely satisfied from the Q-Q plot. 


To visualize the distribution of the residuals, a histogram will be used:
```{r, echo = FALSE, fig.align="center"}
hist(reg1$residuals, main = "Histogram of Residuals", xlab = "Residual Values")
```

From the histogram of the residuals, we can observe that the distribution of the residuals looks symmetric about the mean and bell-shaped, similar to what is seen from a normal distribution. 

To further test for normality of the errors, a Shapiro-Wilk test will be used on the distribution of the residuals. A Shapiro-Wilk test is used to test whether a distribution of data follows a normal distribution.

The null and alternative hypotheses are:
H0: The residuals are normally distributed.
H1: The residuals are not normally distributed.


```{r, echo = FALSE}

W.statistic = round(as.numeric(names(table(shapiro.test(reg1$residuals)$statistic))), 2)
P.value = round(shapiro.test(reg1$residuals)$p.value, 2)

table = cbind(W.statistic, P.value)
colnames(table) = c('W Statistic', 'P-Value')
kable(table, caption = "Shapiro-Wilk Normality Test") %>% kable_styling(bootstrap_options = c("striped", "bordered"), full_width = F)


```


The p-value of 0.13 is greater than the significance level of 0.05, and thus we fail to reject the null hypothesis. Thus, there is no evidence that the distribution of the residuals does not follow a normal distribution. 




# Equal Variances:

Residual vs Fitted Plot:

```{r, echo = FALSE, fig.align="center"}
plot(med_anova, which = c(1))
```


The residuals are pretty evenly spread out along the x-axis, and thus it seems that the equal variance assumption is satisfied.

A Levene's test will be used to further test the equal variance assumption for both variables. 

The null and alternative hypotheses are:
H0: The variances are equal across the various groups.
H1: Not all variances are equal across groups. 


```{r, echo = FALSE}
a = leveneTest(g1_medmath~g1_classtype, data = t_data1)$Df[1]
b = leveneTest(g1_medmath~g1_classtype, data = t_data1)$Df[2]
c = paste(a, b, sep = ", ")
d = round(leveneTest(g1_medmath~g1_classtype, data = t_data1)$F[1], 2)
e = round(leveneTest(g1_medmath~g1_classtype, data = t_data1)$P[1], 2)

table1 = cbind(c, d, e)
colnames(table1) = c('df', 'F Value', 'Pr(>F)')
kable(table1, caption = "Levene Test for Class Type Variable") %>% kable_styling(bootstrap_options = c("striped", "bordered"), full_width = F)

```


```{r, echo = FALSE}
a = leveneTest(g1_medmath~g1_schoolID, data = t_data1)$Df[1]
b = leveneTest(g1_medmath~g1_schoolID, data = t_data1)$Df[2]
c = paste(a, b, sep = ", ")
d = round(leveneTest(g1_medmath~g1_schoolID, data = t_data1)$F[1], 2)
e = round(leveneTest(g1_medmath~g1_schoolID, data = t_data1)$P[1], 2)

table1 = cbind(c, d, e)
colnames(table1) = c('df', 'F Value', 'Pr(>F)')
kable(table1, caption = "Levene Test for School ID Variable") %>% kable_styling(bootstrap_options = c("striped", "bordered"), full_width = F)
```


In the Levene's test for both variables, both p-values are greater than the significance level of 0.05. Thus for each variable, there is no evidence that not all of the variances are equal across treatment groups. Thus, the equal variance assumption is satisfied.  






# Independence:

This experiment is randomized in two ways and represents the best controlled environment to achieve independence of the results. First of all, teachers are randomly assigned to each class type: small, regular, and regular with aide. Second, every student is randomly assigned to each teacher. There is a possibility of the independence assumption not holding completely from the time of randomization to the time that the test scores were recorded. For instance, teachers may share materials, or parents may seek out tutoring for their child. However, given the randomized design of the experiment and the relatively large sample size, we may assume that the results are essentially independent. 



# Outliers:

An outlier will be defined if it has a studentized residual value that is greater than 3 in absolute value.

```{r, echo = FALSE}

a = length(which(studres(reg1)>3 | studres(reg1)< -3))
b = names(which(studres(reg1)>3 | studres(reg1)< -3))

table2 = cbind(a, b)
colnames(table2) = c('Number of Outliers', 'Observation Numbers')
kable(table2, caption = "Outliers") %>% kable_styling(bootstrap_options = c("striped", "bordered"), full_width = F)

```

There is only one outlier in the data, which is observation number 252. Thus, outliers do not pose a problem in our analysis.





